---
title: "'Death Con Three on the Jews': An Analysis of Jewish Life in America Amid Trends of Growing Antisemitism"
author: "Ruty Korotaev"
date: "March 25, 2023"
abstract: "Amid trends of growing antisemitism in the United States and across the world, this report will look at survey data of American Jews between the years 1988 and 2021. The report explored variables related to how Jewish respondents identified themselves, their religious denominations, upbringing information, spousal data, and more. While there were numerous key issues with the GSS dataset, this report demonstrates that there has been a considerable shift in America's Jewish community since 1988, including in how people identify, intermarriage rates, and more."
format: pdf
editor: visual
---

## Introduction

In October 2022, rapper Kanye West went on a Twitter rampage, claiming that he will be going "death con three on the Jewish people," and yet, ascertaining that he cannot be an antisemite because "black people are actually Jew \[sic\]." In the same Tweet, West alludes to the centuries-old belief that the Jewish people have control over the media and are able to remove any person that opposes their collective "agenda." While antisemitic remarks such as these are not new, Kanye West's openness and readiness to target Jewish people online is demonstrative of a larger trend throughout the United States and much of the Western world, which seems to push antisemitism from the margins of society into the mainstream. In this political context, this will use data collected from Jewish respondents participating in the U.S. General Survey (GSS) between 1988 and 2021, in order to understand the state of Jewish life in the U.S., how it has changed over the last few decades, and what indicators contribute to high Jewish affiliation later in life.

For context, the GSS is a national survey conducted periodically to collect data on various aspects of American life, including demographic, social, and economic factors. One important aspect of this survey is the collection of data on religious affiliation, which allows researchers to track the size and characteristics of different religious groups in the United States. Within this framework, the survey also collects data on Jewish Americans, including their religious beliefs and practices, demographic characteristics, education, and their social and economic status.

In addition, there are several other databases that are updated regularly which look at various trends in America's Jewish community. The most notable report is conducted by the Pew Research Center, the last iteration of which was published in May 2021 and revealed several important findings about the characteristics, beliefs, and experiences of this population. One key finding was that the American Jewish population is becoming increasingly diverse, with growing numbers of Jews of colour and Jews who identify as non-binary or transgender. The study also found that younger Jews are less religiously observant than older generations, with higher rates of intermarriage and lower levels of synagogue attendance. Additionally, the study highlighted the importance of Jewish identity to many American Jews, with 75% saying that being Jewish is either very important or somewhat important to them. Finally, the study explored political views among American Jews, finding that a majority identify as Democrats and that there are significant differences in political attitudes between Orthodox and non-Orthodox Jews \[\@citepewresearch\].

While the GSS was not nearly as comprehensive as the Pew Research Study, the findings of this paper seek to corroborate some of the findings that were featured in Pew's research, to gain a more fulsome understanding of Jewish life in the United States.

## Data Introduction

This paper uses data collected from the US General Survey, and visualizes data related to Jewish identity and practice. As part of this, the main variables that were used in this project include year, age, respondent ID, what denomination of Judaism they ascribe to, whether they were raised as Jews and if they still identify as Jews, as well as marital information, including whether their spouses were Jewish and if they were also raised as Jews. This paper uses R \[\@citeR\] to analyze a dataset from the US General Survey which shares data on the population flow within Toronto's shelter system. In addition, several R packages were included in the project, including "tidyverse" \[\@citetidyverse\], "dplyr" \[\@citedplyr\], "kableExtra" \[\@citekableExtra\], and "knitr" \[\@citeknitr\].

## Jewish Denominations: Overview

Within Judaism, there are numerous denominations that most members of the community tend to identify with. This section will show visualizations for the data collected on denominations, focusing specifically on those identifying in the three main groups, including Orthodox, Conservative, and Reform. @fig-reform-dem visualizes the amount of respondents who identified as Reform Jews, while @fig-conservative-dem shows who identified as Conservative, and @fig-orthodox-dem depicts the Orthodox respondents in the dataset.

```{r}
1 + 1
#| echo: false
#| message: false
#| label: fig- 
#| tbl-cap: "Amount of Jewish respondents between being raised in Reform Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 3) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Reform"
  )

```

```{r}
1 + 1
#| echo: false
#| message: false
#| label: fig-reform-dem
#| tbl-cap: "Amount of Jewish respondents identifying with Reform Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 3) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Reform"
  )
```

```{r}
#| echo: false
#| message: false
#| label: fig-orthodox-dem
#| fig-cap: "Amount of Jewish respondents identifying with Orthodox Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 1) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Orthodox"
  )

```

```{r}
#| echo: false
#| message: false
#| label: fig-conservative-dem
#| fig-cap: "Amount of Jewish respondents identifying with Conservative Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 2) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Conservative"
  )
```

In addition to just the demographic data, this report will also look at how one's denomination growing up can impact how a respondent identifies later on in life. @fig-upbringing looks at this correlation and how being raised with a denomination impacts identity later on.

```{r}
#| echo: false
#| message: false
#| label: fig-upbringing
#| fig-cap: "Correlation between upbringing and Jewish identity as an adult"
library(tidyr)
library(ggplot2)

library(tidyr)
library(ggplot2)

clean_respondent_data %>%
  pivot_longer(cols = c("jew16", "jew"), names_to = "upbringing", values_to = "value") %>%
  ggplot(aes(x = variable, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_classic() +
  labs(x = "Year", y = "Jewish Affiliation", fill = "Upbringing")

```

## Marital Data

Among the chief findings in the Pew 2020 report on Jewish life in America was that intermarriage (marriages between Jews and those not identifying as Jews) was lower among Orthodox Jews, and higher among those affiliated with the Reform movement. This section seeks to corroborate and see whether this was also a finding in the GSS data. In @fig-conservative-marriage, @fig-reform-marriage, and @fig-orthodox-marriage, we look at how members of the three main denominations have chosen their life partners and what denomination they ascribe to.

```{r}
#| echo: false
#| message: false
#| label: fig-reform-marriage
#| fig-cap: "Marriage statistics in Reform Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 3, spjew %in% c(1, 2, 3, 4, 5)) %>%
  ggplot(aes(x = year, y = jew, 
             color = spjew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Marriage"
  )

```

```{r}
#| echo: false
#| message: false
#| label: fig-orthodox-marriage
#| fig-cap: "Marriage statistics in Orthodox Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 1, spjew %in% c(1, 2, 3, 4, 5)) %>%
  ggplot(aes(x = year, y = jew, 
             color = spjew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Marriage"
  )


```

```{r}
#| echo: false
#| message: false
#| label: fig-conservative-marriage
#| fig-cap: "Marriage statistics in Conservative Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 2, spjew %in% c(1, 2, 3, 4, 5)) %>%
  ggplot(aes(x = year, y = jew, 
             color = spjew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Marriage"
  )

```

## Results

The results of the above visualizations demonstrate that Jewish life in the United States has diverse and dynamic, which contradicts the often-used antisemitic trope that tends to homogenize the entire community. In the Jewish demographic data section, it can be noted that the majority of respondents identify as either Reform, Orthodox, or Conservative, and that the upbringing can play a key role in the ways people identify later on in life.

Further, in the marital data section, we can see how the various denominations choose their spouses is often based on their own identities and religious affiliations. Much in line with the Pew research report, there was a much lower intermarriage rate amount Orthodox Jews, who typically married those who were also part of the Orthodox community. At the same time, intermarriage was higher among Conservative and Reform Jews, who typically are not as strict in their religious practice as Orthodox Jews.

## Discussion: Data Findings

The data demonstrates that there have been several key changes within American Jewish life since the late 1980s, when this survey data first started to be collected.

## Discussion: Survey Limitations

This survey could be improved in a variety of ways, most notably in the fact that there is a great deal of missing data, with almost entire variables having blank columns. While it is made clear that this survey was not meant to specifically target the Jewish community, in attempting to include people from all religious groups and denominations, it effectively leaves the dataset with little information at all. As such, in attempting to understand the various religious communites throughout the United States, it may be helpful to have more targeted, specialized surveys in order to better understand these demographics.

Further, while the survey seems to do a good job in identifying factors such as upbringing and spousal choices as being key aspects of Jewish life, it fails to represent the many nuances that exist within the Jewish community. This starts with the denominations themselves -- GSS only looked at three main denominations, but failed to include the many other ones that exist in the community, including Haredi, Reconstructionist, and Modern Orthodox, as well as omitting the many Jews who identify as culturally Jewish but are non-practicing. It also fails to ask the respondents about their own Jewish cultural background, including whether they are Ashkenazi, Sephardi, or Mizrahi, which is also a key part of many Jewish people's identities and practices. Further, GSS does not include any data on immigrant Jews who make up a growing amount of percentage of Jews living in the United States in recent years, including those hailing from the former Soviet Union, Israel and the Middle East, Europe, North Africa, and other regions. Overall, the GSS dataset was lacking in both nuance and responses, making it difficult to work with and to come to proper conclusions.
