---
title: "'Death Con Three on the Jews': An Analysis of Jewish Life in America Amid Trends of Growing Antisemitism"
author: "Ruty Korotaev"
date: "March 25, 2023"
abstract: "Amid trends of growing antisemitism in the United States and across the world, this report will look at survey data of American Jews between the years 1988 and 2021. The report explored variables related to how Jewish respondents identified themselves, their religious denominations, upbringing information, spousal data, and more. This data demonstrates that since 1988, there has been a considerable ____. "
format: pdf
editor: visual
---

## Introduction

In October 2022, rapper Kanye West went on a Twitter rampage, claiming that he will be going "death con three on the Jewish people," and yet, ascertaining that he cannot be an antisemite because "black people are actually Jew \[sic\]." In the same Tweet, West alludes to the centuries-old belief that the Jewish people have control over the media and are able to remove any person that opposes their collective "agenda." While antisemitic remarks such as these are not new, Kanye West's openness and readiness to target Jewish people online is demonstrative of a larger trend throughout the United States and much of the Western world, which seems to push antisemitism from the margins of society into the mainstream. In this political context, this will use data collected from Jewish respondents participating in the U.S. General Survey (GSS) between 1988 and 2021, in order to understand the state of Jewish life in the U.S., how it has changed over the last few decades, and what indicators contribute to high Jewish affiliation later in life.

For context, the GSS is a national survey conducted periodically to collect data on various aspects of American life, including demographic, social, and economic factors. One important aspect of this survey is the collection of data on religious affiliation, which allows researchers to track the size and characteristics of different religious groups in the United States. Within this framework, the survey also collects data on Jewish Americans, including their religious beliefs and practices, demographic characteristics, education, and their social and economic status.

In addition, there are several other databases that are updated regularly which look at various trends in America's Jewish community. The most notable report is conducted by the Pew Research Center, the last iteration of which was published in May 2021 and revealed several important findings about the characteristics, beliefs, and experiences of this population. One key finding was that the American Jewish population is becoming increasingly diverse, with growing numbers of Jews of colour and Jews who identify as non-binary or transgender. The study also found that younger Jews are less religiously observant than older generations, with higher rates of intermarriage and lower levels of synagogue attendance. Additionally, the study highlighted the importance of Jewish identity to many American Jews, with 75% saying that being Jewish is either very important or somewhat important to them. Finally, the study explored political views among American Jews, finding that a majority identify as Democrats and that there are significant differences in political attitudes between Orthodox and non-Orthodox Jews \[\@citepewresearch\].

While the GSS was not nearly as comprehensive as the Pew Research Study, the findings of this paper seek to corroborate some of the findings that were featured in Pew's research, to gain a more fulsome understanding of Jewish life in the United States.

## Data Introduction

This paper uses data collected from the US General Survey, and visualizes data related to Jewish identity and practice. As part of this, the main variables that were used in this project include year, age, respondent ID, what denomination of Judaism they ascribe to, whether they were raised as Jews and if they still identify as Jews, as well as marital information, including whether their spouses were Jewish and if they were also raised as Jews. This paper uses R \[\@citeR\] to analyze a dataset from the US General Survey which shares data on the population flow within Toronto's shelter system. In addition, several R packages were included in the project, including "tidyverse" \[\@citetidyverse\], "dplyr" \[\@citedplyr\], "kableExtra" \[\@citekableExtra\], and "knitr" \[\@citeknitr\].

## Jewish Denominations: Overview

Within Judaism, there are numerous denominations that most members of the community tend to identify with. This section will show visualizations for the data collected on denominations, focusing specifically on those identifying in the three main groups, including Orthodox, Conservative, and Reform. @fig-reform-dem visualizes the amount of respondents who identified as Reform Jews, while @fig-conservative-dem shows who identified as Conservative, and @fig-orthodox-dem depicts the Orthodox respondents in the dataset.

```{r}
1 + 1
#| echo: false
#| message: false
#| label: fig- 
#| tbl-cap: "Amount of Jewish respondents between being raised in Reform Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 3) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Reform"
  )

```

```{r}
1 + 1
#| echo: false
#| message: false
#| label: fig-reform-dem
#| tbl-cap: "Amount of Jewish respondents identifying with Reform Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 3) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Reform"
  )
```

```{r}
#| echo: false
#| message: false
#| label: fig-orthodox-dem
#| fig-cap: "Amount of Jewish respondents identifying with Orthodox Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 1) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Orthodox"
  )

```

```{r}
#| echo: false
#| message: false
#| label: fig-conservative-dem
#| fig-cap: "Amount of Jewish respondents identifying with Conservative Judaism"

#load library 
library(tidyverse)
library(dplyr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data %>% 
  filter(jew == 2) %>%
  ggplot(aes(x = year, y = jew, 
             color = jew)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(x = "Year", y = "Respondents Identifying as Conservative"
  )

```

In considering the amount of

## Results

Based on the above visualizations, it is clear that Jewish life and identity within the United States changes based on the real or percieved threats of antisemitism.

Additionally, as the data shows, the demographic that is growing the most -- Orthodox and immigrant Jews -- is frequently left out of the conversation surrounding Jewish life in the United States, with many of the Jewish voices being of those who are at two to three generations living in the US, most commonly those who are of the reform denomination.

## Discussion

The data demonstrates that over the last 20 years....

## Appendix: Supplemental Survey

This survey could be improved in a variety of ways. First, given the fact that there is a massive amount of missing data in the survey, it seems to indicate that not very many Jews were actually interviewed for this survey. I believe that the survey should have focused on people who identify as Jews, the way that the Pew Research Centre does it, in order to avoid having so much missing data.

The questions for the survey could have been more nuanced or detailed.
