---
title: "'Death Con Three on the Jews': An Analysis of Jewish Life in America Amid Trends of Growing Antisemitism"
author: "Ruty Korotaev"
date: "March 25, 2023"
abstract: "Amid trends of growing antisemitism in the United States and across much of the Western world, this report will look at survey data of American Jews between the years 2000 and 2021. The report explored variables related to how Jewish respondents identified, how closely affiliated they were to their identities, as well as demographic data related to their specific Jewish denominations, spousal religion, and upbringing. This data demonstrates that over the last 20 years_____"
format: pdf
editor: visual
---

## Introduction

This paper focuses on analyzing Jewish life and identity in America from 2000 to 2021, based on data collected from the United States' General Social Survey.

The US General Survey is a national survey conducted periodically to collect data on various aspects of American life, including demographic, social, and economic factors. One important aspect of this survey is the collection of data on religious affiliation, which allows researchers to track the size and characteristics of different religious groups in the United States. Within this framework, the survey also collects data on Jewish Americans, including their religious beliefs and practices, demographic characteristics, and social and economic status. This data can be used by policymakers, academics, and community leaders to better understand and address the needs and experiences of Jewish Americans, as well as to promote interfaith understanding and cooperation.

There a several other databases that are updated annually, which look at various trends in America's Jewish community. The most notable one is conducted by the Pew Research Center, the last iteration of which was published in May 2021 and revealed several important findings about the characteristics, beliefs, and experiences of this population. One key finding was that the American Jewish population is becoming increasingly diverse, with growing numbers of Jews of color and Jews who identify as non-binary or transgender. The study also found that younger Jews are less religiously observant than older generations, with higher rates of intermarriage and lower levels of synagogue attendance. Additionally, the study highlighted the importance of Jewish identity to many American Jews, with 75% saying that being Jewish is either very important or somewhat important to them. Finally, the study explored political views among American Jews, finding that a majority identify as Democrats and that there are significant differences in political attitudes between Orthodox and non-Orthodox Jews. (ADD CITATIONS)

## Data

This paper uses data collected from the US General Survey, and visualizes data related to Jewish identity and practise. As part of this, the main variables that were used in this project include year, age, respondent ID, what denomination of Judaism they ascribe to, whether they were raised as Jews and if they still identify as Jews, as well as marital information, including whether their spouses were Jewish and if they were also raised as Jews. This paper uses R \[\@citeR\] to analyze a dataset from the US General Survey which shares data on the population flow within Toronto's shelter system. In addition, several R packages were included in the project, including "tidyverse" \[\@citetidyverse\], "dplyr" \[\@citedplyr\], "kableExtra" \[\@citekableExtra\], and "knitr" \[\@citeknitr\].

```{r}
1 + 1
#| echo: false
#| message: false
#| label: fig- 
#| tbl-cap: "Amount of Jewish respondents between being raised in Reform Judaism"

#load library 
library(tidyverse)
library(knitr)

#Load dataset
clean_respondent_data <- 
  read_csv(here::here("outputs/cleaned_respondent_data.csv"))

#Graph data
clean_respondent_data |>
  ggplot(aes(x = year, 
             y = relig, 
             color = jew16)) +
  geom_line() +
  geom_point()+
  theme_classic() +
  labs(
    x = "Year",
    y = "Jewish identity and raising"
  )

```

You can add options to executable code like this

```{r}
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: "Jewish denominations most common in the United States"


```

## Results

Based on the above visualizations, it is clear that Jewish life and identity within the United States changes based on the real or percieved threats of antisemitism.

Additionally, as the data shows, the demographic that is growing the most -- Orthodox and immigrant Jews -- is frequently left out of the conversation surrounding Jewish life in the United States, with many of the Jewish voices being of those who are at two to three generations living in the US, most commonly those who are of the reform denomination.

## Discussion

The data demonstrates that over the last 20 years....

## Appendix: Supplemental Survey

This survey could be improved in a variety of ways. First, given the fact that there is a massive amount of missing data in the survey, it seems to indicate that not very many Jews were actually interviewed for this survey. I believe that the survey should have focused on people who identify as Jews, the way that the Pew Research Centre does it, in order to avoid having so much missing data.

The questions for the survey could have been more nuanced or detailed.
